---
title: "plots try out"
format: html
editor: visual
---

prøver endnu en gang

```{r}

# Loading libraries
#| echo: false
library(tidyverse)
library(dplyr)
library(scales)

# Defining the different locataions and their population in 2021

location <- c("Argentina", "Austria", "Belgium", "Bulgaria", "Croatia","Cyprus", "Czechia", "Denmark", "Ecuador", "Estonia", "Finland", "France", "Germany", "Hong Kong","Hungary","Iceland","Ireland","Italy","Japan","Latvia","Liechtenstein","Lithuania","Luxembourg","Malta","Nepal","Netherlands","Norway","Peru","Poland","Portugal","Romania","Slovakia","Slovenia","South Africa","South Korea","Spain","Sweden","Switzerland","Ukraine", "United States", "Uruguay", "European Union")

# And their populations in 2021 

population <- c(48.81,8.956,11.59,6.878,3.879,1.244,10.51,5.857,17.8,1.331,5.541,67.76,83.2,7.413,9.71,0.372520,5.033,59.13,125.7,1.884,0.039039,2.801,0.640064,0.518536,30.03,17.53,5.408,33.72,37.75,10.36,19.12,5.447,2.108,59.39,51.77,47.42,10.42,8.705,43.85,332,3.426,446.9)

# Creating popualtion tibble: 

population_data_tibble <- tibble(location,population)
```

```{r}
# Checking austria correct answer summary
363548+1585063+3682+1588222+14584985
```

```{r}
Vaccinated_by_country_all_vaccines <- manufacturer_data %>%
  filter(date == '2022-03-25') %>%  # Filter by specific date
  group_by(location) %>%  # Group by location
  summarise(Total_vaccinations = sum(total_vaccinations, na.rm = TRUE))  # Summarize total vaccinations
```

This is the amount of vaccines done within the period. Most people, if given the full dosis, have recieved 2 vaccines during the period. Thus, the number will be divided by 2, to recieve the amount of people having recieved a full dosis.

```{r}
Vaccinated_by_country_all_vaccines <- Vaccinated_by_country_all_vaccines |> 
  mutate(Vaccinated_fully=Total_vaccinations/2)
```

We have made a tibble with the countries and populations for 2021. We want to join this onto the `Vaccinated_by_country_all_vaccines`

```{r}
Clean_table <- Vaccinated_by_country_all_vaccines |> 
  left_join(population_data_tibble,by="location") |> 
  mutate(Vaccinated_fully_in_mio=Vaccinated_fully/1e6)

Clean_table
```

Now we can finally found the percentage of how much of the population was vaccinated by the end of the period, at the date: 2022-03-25.

```{r}
Clean_table <- Clean_table |> 
  mutate(percentage_vaccinated=(Vaccinated_fully_in_mio/population)*100)

Clean_table
```

now we can make the world map:

```{r}
library(maps)
world_map <- map_data("world")

# Now we merge the world map and the data we just did: 

world_map_vaccinated <- world_map |> 
  left_join(Clean_table,by=c("region"="location"))

Clean_table

```

```{r}
ggplot(world_map_vaccinated, aes(x = long, y = lat, group = group, fill = percentage_vaccinated)) +
  geom_polygon(color = "black", size = 0.2) +
  scale_fill_gradient(low = "lightblue", high = "darkblue", na.value = "grey50", name = "Vaccination %") +
  theme_minimal() +
  labs(
    title = "Vaccination Percentage by Country",
    subtitle = "Countries color-coded by vaccination percentage. Grey countries not included.",
    caption = "Data Source: Kaggle",
    x = "hello",
    y = "hello"
    
  ) +
  theme(
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank(),
    legend.position = "bottom"
  )
```

From this we can see which countries have the highest percentage of the population vaccinated by 2022-03-25.
