---
title: "03_aug_data"
author: "Group 02"
format: html
editor: visual
---

## Data augmentation

The manufacturer data and vaccine data contains a lot of the same information, the main differences are size. Therefore we sometimes we make trial runs for some of the code with that data set, but our main code consist of the vaccine data.

```{r}
#| echo: false

# Loading libraries
library(tidyverse)
library(dplyr)
```

We will join the vaccine data and the world population data and making some new variables. But first we load the clean data:

```{r}
#| echo: false
# Loading data
clean_manufacturer <- read_tsv(file = "../data/02_manufacturer_cleaned.tsv")
clean_vaccines <- read_tsv(file = "../data/02_vaccines_cleaned.tsv")
clean_world_population <- read_tsv(file = "../data/02_world_population_cleaned.tsv")
```

We make a simple change to vaccines:

```{r}
# Adding variable to vaccines
aug_vaccine <- clean_vaccines |> 
    mutate( # Flaging if there is missing some information
    missing_total_vaccinations = is.na(total_vaccinations),
    missing_daily_vaccinations = is.na(daily_vaccinations)) |> 
  rename(`Country` = country) # Changing name
```

Now let make a new population estimate for 2021 by using the growth rate and the 2020 population:

```{r}
# Creating new variable 
options(scipen = 999) # Removing scientific notation

aug_world_population <- clean_world_population |> 
  mutate("2021 Population" = round(`2020 Population`*`Growth Rate`,digits = 0)) |> 
  # Rounding the digits 
  select(`Country/Territory`, # Selecting relevant columns
         `Continent`,
         `2022 Population`,
         `2021 Population`, 
         `2020 Population`) |> 
  rename(`Country` = `Country/Territory`)
```

Now we join the data:

```{r}
# Joining data
## Vaccine and world population

## Manufacture and world population
manufacturer_world_pop <- clean_manufacturer |> 
  rename(`Country` = location) |> # Renaming location
  left_join(aug_world_population,
            by = "Country")
```

Further manipulation, first of with manufacturer and world population data:

```{r}
# Finding the total amount of vaccination pr country
manufacturer_world_pop <- manufacturer_world_pop |> 
  group_by(Country) |> # Grouping by country and taking to sum of all vaccinations
  arrange(date) |> # Making the tibble cronological 
  mutate(`Vaccinations pr country` = cumsum(total_vaccinations)) # Cumulative sum

# Finding the total amount of people vaccinated pr country - 2 vaccination pr person
manufacturer_world_pop <- manufacturer_world_pop |>
  mutate(`Vaccinated people pr country` = round(`Vaccinations pr country`/2,0))

```

We have decided that we want to pivot the populations so they become longer, and then we can make a data set/database:

```{r}
# Pivoting the population in seperate variable 
population_data <-  manufacturer_world_pop |> 
  pivot_longer(cols = starts_with("20"), # Columns starting with 20 pivot longer
               names_to = "Year", # New variable name "Year"
               values_to = "Population") |> # Variable for value added, named "Population"
  mutate(Year = as.numeric(str_remove(Year, " Population"))) # Remove the word "Population"

# Extracting Year informantion from the date and making a variable showing the year
manufacturer_world_pop <- manufacturer_world_pop |> 
  mutate(Year = as.numeric(format(date, "%Y")))

# Joining the population data with the new manufacturer world population data
manufacturer_world_pop <- manufacturer_world_pop |> 
  left_join(population_data, c("Country", "Year")) # Joining based on country and year


```
