---
title: "04_analysis" 
format: html 
editor: visual 
---

## Analysis and plots - world plots

First of we'd like to make a heat map over the world when taking vaccine amount in to account.

We load necessary libraries:

```{r}
library(tidyverse)
library(scales)
```

Load the data:

```{r}
#Loading data
manufac_world_pop <- read_tsv("../data/03_manufacturer_world_population_aug.tsv")

```

```{r}
# Filtering data by country for easier overview 

manufac_world_pop <- manufac_world_pop |> 
  arrange(by_group=Country)

  
```

### Plotting the actual plot

importing another data set so we can make a heat map. We will combine the data here but not make a new data file because of it being to big to load.

```{r}
# Loading the maps package

library(maps)
world_map <- map_data("world")

# Now we merge the world map and the data we just did: 
world_map_vaccinated <- world_map |> 
  left_join(manufac_world_pop, by = c("region" = "Country"), relationship = "many-to-many")

```

Plotting heat map for different dates through the data set

```{r}
# Plotting with ggplot
## First month of the vaccines
p1 <- world_map_vaccinated |> 
  filter(date == as.Date("2020-12-31")) |> 
  ggplot(aes(x = long, y = lat, group = group, fill = `Percentage Vaccinated`)) +
    geom_polygon(color = "black", size = 0.2) +
    scale_fill_gradient(low = "lightblue", 
                        high = "darkblue", 
                        na.value = "grey50", 
                        name = "Vaccination %") +
    theme_minimal() +
    labs(
      title = "Vaccination Percentage by Country, first month (2020-12-31)",
      subtitle = "Countries color-coded by vaccination percentage. Grey countries not included.",
      caption = "Data Source: Kaggle") +
    theme(
      axis.text = element_blank(),
      axis.ticks = element_blank(),
      panel.grid = element_blank(),
      legend.position = "bottom"
    ) 
p1
```

```{r}
# Plotting with ggplot
p2 <- world_map_vaccinated |> 
  filter(date == as.Date("2021-05-04")) |> 
  ggplot(aes(x = long, y = lat, group = group, fill = `Percentage Vaccinated`)) +
    geom_polygon(color = "black", size = 0.2) +
    scale_fill_gradient(low = "lightblue", 
                        high = "darkblue", 
                        na.value = "grey50", 
                        name = "Vaccination %") +
    theme_minimal() +
    labs(
      title = "Vaccination Percentage by Country, first month (2021-05-04)",
      subtitle = "Countries color-coded by vaccination percentage. Grey countries not included.",
      caption = "Data Source: Kaggle") +
    theme(
      axis.text = element_blank(),
      axis.ticks = element_blank(),
      panel.grid = element_blank(),
      legend.position = "bottom"
    ) 
p2
```

```{r}
# Plotting with ggplot
p3 <- world_map_vaccinated |> 
  filter(date == as.Date("2022-02-13")) |> 
  ggplot(aes(x = long, y = lat, group = group, fill = `Percentage Vaccinated`)) +
    geom_polygon(color = "black", size = 0.2) +
    scale_fill_gradient(low = "lightblue", 
                        high = "darkblue", 
                        na.value = "grey50", 
                        name = "Vaccination %") +
    theme_minimal() +
    labs(
      title = "Vaccination Percentage by Country, first month (2022-02-13)",
      subtitle = "Countries color-coded by vaccination percentage. Grey countries not included.",
      caption = "Data Source: Kaggle") +
    theme(
      axis.text = element_blank(),
      axis.ticks = element_blank(),
      panel.grid = element_blank(),
      legend.position = "bottom"
    ) 
p3
```

```{r}
# Plotting with ggplot
p4 <- world_map_vaccinated |> 
  ggplot(aes(x = long, y = lat, group = group, fill = `Percentage Vaccinated`)) +
    geom_polygon(color = "black", size = 0.2) +
    scale_fill_gradient(low = "lightblue", 
                        high = "darkblue", 
                        na.value = "grey50", 
                        name = "Vaccination %") +
    theme_minimal() +
    labs(
      title = "Vaccination Percentage by Country, first month (2020-12-31)",
      subtitle = "Countries color-coded by vaccination percentage. Grey countries not included.",
      caption = "Data Source: Kaggle") +
    theme(
      axis.text = element_blank(),
      axis.ticks = element_blank(),
      panel.grid = element_blank(),
      legend.position = "bottom"
    ) 
```

```{r}
p1
p2
p3
p4
```
