---
title: "04_analysis"
format: html
editor: visual
---

## Analysis and plots

<<<<<<< HEAD
```         
=======
First of we'd like to make a heat map over the world when taking vaccine amount in to account.

We load necessary libraries:

```{r}
# Libraries
library(tidyverse)
library(scales)
```

Load the data:

```{r}
# Loading data
manufac_wolrd_pop <- read_tsv("../data/03_manufacturer_world_population_aug.tsv")
```

bla bla

```{r}
# Creating a heat map based on the percentage vaccinated pr country in the data set
ggplot(manufac_wolrd_pop, aes(x = long, y = lat, group = group, fill = `Percentage Vaccinated`)) +
  geom_polygon(color = "black", size = 0.2) +
  scale_fill_gradient(low = "lightblue", high = "darkblue", na.value = "grey50", name = "Vaccination %") +
  theme_minimal() +
  labs(
    title = "Vaccination Percentage by Country",
    subtitle = "Countries color-coded by vaccination percentage. Grey countries not included.",
    caption = "Data Source: Kaggle",
    x = "hello",
    y = "hello"
    
  ) +
  theme(
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank(),
    legend.position = "bottom"
  )
>>>>>>> 8f5d76b4d1ea8f8908d8a3f7d60c418cecaa0e5e
```
