---
title: "07_analysis_manufacturer_progression_plot.qmd"
format:
  html:
    embed-resources: true
editor: visual
---

## Progression of the different vaccines based on countries

We would like to see the differences of countries choices of vaccine and see if some vaccines were especially popular. We can also make some general comparisons between the different continents.

First we load libraries:

```{r}
#| echo: false 
library(tidyverse)
```

We load data:

```{r}
#| echo: false 
data <- read_tsv("../data/03_manufacturer_world_population_aug.tsv")
```

Now we can make a plot based on the dates on the x-axis and amount of each vaccine on the y-axis:

```{r}
# Plotting: 
## Denmark the first 4 months of 2021
p1 <- data |> 
  filter(Country == "Denmark",
         date >= "2021-01-01" & date <= "2021-04-01") |> 
  ggplot(aes(x = date,
             y = total_vaccinations,
             color = vaccine)) +
  geom_line(aes(group = vaccine)) 

p1
## Denmark 
p2 <- data |> 
  filter(Country == "Denmark") |> 
  ggplot(aes(x = date,
             y = total_vaccinations,
             color = vaccine)) +
  geom_line(aes(group = vaccine)) 

p2
## Argentina 
p3 <- data |> 
  filter(Country == "Argentina") |> 
  ggplot(aes(x = date,
             y = total_vaccinations,
             color = vaccine)) +
  geom_line(aes(group = vaccine)) 
p3
## Germany the first 4 months of 2021
p4 <- data |> 
  filter(Country == "Germany",
         date >= "2021-01-01" & date <= "2021-04-01") |> 
  ggplot(aes(x = date,
             y = total_vaccinations,
             color = vaccine)) +
  geom_line(aes(group = vaccine)) 
p4
## Germany
p5 <- data |> 
  filter(Country == "Germany") |> 
  ggplot(aes(x = date,
             y = total_vaccinations,
             color = vaccine)) +
  geom_line(aes(group = vaccine)) 
p5
```
