---
title: "10_vaccine_list_analysis"
format:
  html:
    embed-resources: true
editor: visual
---

## Vaccines Usage Analysis

We wish to form a list showcasing the most used vaccines, and the locations they have been mostly used.

```{r}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(knitr)
```

```{r}
# We load the cleaned and augmented data
vaccines_cleaned <- read_tsv("../data/02_vaccines_cleaned.tsv")
manufacturer_cleaned <- read_tsv("../data/02_manufacturer_cleaned.tsv")

# then we view the structure of the loaded data
glimpse(vaccines_cleaned)
glimpse(manufacturer_cleaned)


```

## Data Analysis

### Most used vaccines

We now find the most used vaccines

```{r}
# Let us calculate the total usage of each vaccine
most_used_vaccines <- manufacturer_cleaned |>
  group_by(vaccine) |>
  summarise(Total_Usage = sum(total_vaccinations, na.rm = TRUE)) |>
  arrange(desc(Total_Usage))

most_used_vaccines
```

### Countries in use

We can now identify the country with the highest usage for each vaccine

```{r}
vaccine_by_country <- manufacturer_cleaned |>
  group_by(vaccine, location) |>
  summarise(Total_Usage = sum(total_vaccinations, na.rm = TRUE), .groups = "drop") |> # We now drop grouping after summarise
  arrange(desc(Total_Usage)) |>
  group_by(vaccine) |> # We need to re-group, but only for vaccines so we get the correct order
  slice_max(Total_Usage, n = 1) |> 
  ungroup() |>
  arrange(desc(Total_Usage)) |>
  select(vaccine, location, Total_Usage) # Here we select the order of columns

# View the result
vaccine_by_country

```

### Nested List

We see this, but let us now expand the list. Seeing for each vaccines, the countries and the total vaccinations done by said vaccine in each country.

```{r}
vaccine_country_list <- manufacturer_cleaned |>
  group_by(vaccine, location) |>
  summarise(Total_Usage = sum(total_vaccinations, na.rm = TRUE), .groups = "drop") |>
  arrange(vaccine, desc(Total_Usage)) |> # We here arrange by vaccine and usage for each vaccine
  group_by(vaccine) |>
  nest(data = c(location, Total_Usage)) # Here we nest the country data for each vaccine

vaccine_country_list

```

```{r}
expanded_data <- vaccine_country_list |>
  mutate(data = map_chr(data, ~paste(.x$location, ":", .x$Total_Usage, collapse = "; ")))

kable(expanded_data, format = "html", longtable = TRUE, escape = FALSE)

```

```{r}
expanded_data
```
